Article figures

### 2020-11-04 src/article_figures.Rmd
Polished article plots

  plot <- result %>% 
    ggplot() +     
    geom_vline(aes(xintercept=0), color="lightgray") +
    ggstance::geom_pointrangeh(aes(y=Pretty, x=mean, xmin=low, xmax=high, color=gender),
                               position=position_dodge2(width=0.8, padding=0.1), size=0.25) + # ggstance is required to make legend keys horizontal
    geom_rect(data=combined_variables %>% filter(stripe==1), 
              mapping=aes(ymax = as.numeric(Pretty) + 0.5,
                  ymin = as.numeric(Pretty) - 0.5),
                  #xmin = -Inf,
                  #xmax = Inf,
                  fill = "gray", xmin=-Inf, xmax=Inf, alpha = 0.1, show.legend = FALSE, colour=NA) +
    labs(title="", #title="Effects sizes of variables on Hb prediction",
         x="Regression coefficient", y=NULL, colour="Sex") +
    scale_colour_discrete(drop=FALSE, breaks = c("male", "female")) + 
    #guides(colour = guide_legend(reverse = TRUE)) + # Reverse the order of keys in the legend
    #theme_gray(base_size = base_size) + theme(legend.position = "bottom", legend.direction = "horizontal")
    theme_classic(base_size = base_size) + theme(legend.position = "bottom", legend.direction = "horizontal")
  return(list(plot=plot, cis=cis))
}

@@

  varimp.plot <- var_imp %>% 
    ggplot() + 
    ggstance::geom_pointrangeh(aes(y=Pretty, x=Importance, xmin=0, xmax=Importance, colour = gender), # ggstance is required to make legend keys horizontal
                               position=position_dodge2(width=0.8, padding=0.1), size=0.25, inherit.aes = FALSE) +   
                               #colour = hue_pal()(3)[2], size=0.25) +
      geom_rect(data=combined_variables %>% filter(stripe==1),
              mapping=aes(ymax = as.numeric(Pretty) + 0.5,
                    ymin = as.numeric(Pretty) - 0.5),
                    #xmin = -Inf,
                    #xmax = Inf,
                    #fill = stripe), 
              xmin=-Inf, xmax=Inf, alpha = 0.1, colour=NA, fill="gray", show.legend = FALSE, inherit.aes = FALSE) +
    labs(y=NULL, colour="Sex") + 
    guides(y="none") +
    scale_colour_discrete(drop=FALSE, breaks=c("both")) +
    #theme_gray(base_size = base_size)
    theme_classic(base_size = base_size) + theme(legend.position = "bottom", legend.direction = "horizontal")

@@

```{r}
if (save_figs) {
    filename <- paste(fig_path, "effect_size_importance.pdf", sep="/")
} else {
    filename <- NULL
}
create_forest_importance_plot(male_posterior, female_posterior, variables, filename=filename)
```
## Alternative effect sizes and importances of variables

@@

  var_imp <- varimp %>%
    mutate(Pretty=fct_rev(Pretty),
           mean=Importance,
           low=0,
           high=Importance)

  final <- bind_rows(`Regression coefficient`=result, Importance=var_imp, .id="method") %>%
    mutate(method=factor(method, levels=c("Regression coefficient", "Importance")))

  plot <- final %>% 
    ggplot() +     
    geom_vline(aes(xintercept=0), color="lightgray") +
    ggstance::geom_pointrangeh(aes(y=Pretty, x=mean, xmin=low, xmax=high, color=gender),
                               position=position_dodge2(width=0.8, padding=0.1), size=0.25) + # ggstance is required to make legend keys horizontal
    geom_rect(data=combined_variables %>% filter(stripe==1), 
              mapping=aes(ymax = as.numeric(Pretty) + 0.5,
                          ymin = as.numeric(Pretty) - 0.5),
              fill = "gray", xmin=-Inf, xmax=Inf, alpha = 0.1, show.legend = FALSE, colour=NA) +
    labs(title=NULL,
         x=NULL, y=NULL, colour="Sex") +
    scale_colour_discrete(drop=FALSE, breaks = c("male", "female", "both")) + 
    facet_wrap("method", scales="free_x") +
    theme_classic(base_size = base_size) + 
    theme(legend.position = "bottom", legend.direction = "horizontal", strip.background = element_rect(fill = NULL, color = "white", size = 1))

#    ggstance::geom_pointrangeh(aes(y=Pretty, x=Importance, xmin=0, xmax=Importance, colour = gender), # ggstance is required to make legend keys horizontal
#                               position=position_dodge2(width=0.8, padding=0.1), size=0.25, inherit.aes = FALSE) +

    #inherit.aes = FALSE) +
#    guides(y="none") +

@@

```{r}
g <- alternative_create_forest_importance_plot(male_posterior, female_posterior, variables, filename=filename)

if (save_figs)
  ggsave(filename="alternative_effect_size_importance.pdf", title="Performance forest plot", path=fig_path, plot=g, dpi=600, units="mm", width=180)
g
```

@@

get_threshold_df <- function() {
  t6 <- tibble(
      Id = factor(ids),
      type = factor("Threshold"),
      value = c(0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.2, 0.1, 0.4, 0.6))
  t12 <- tibble(
      Id = factor(ids),
      type = factor("Threshold"),
      value = c(0.2, 0.2, 0.1, 0.1, 0.1, 0.2, 0.3, 0.3, 0.8, 0.8)
  )
  t <- bind_rows(t6, t12)
  t <- t %>%
    mutate(low=value, 
           high=value,
           sex=factor(str_split_fixed(Id, "-", 3)[,2], levels=c("female", "both", "male")),
           model=factor(str_split_fixed(Id, "-", 3)[,3]))
  return(t)
}

@@

  thresholds <- get_threshold_df()
  df <- bind_rows(df, thresholds)

  # Below we use a dummy table and geom_blank to have the same xlimits for E6 and E12 panels
  xrange <- as.numeric(df %>% filter(type %in% c("E6", "E12")) %>% summarise(low=min(low), high=max(high))) # range of x-axis for E6 and E12
  xrange <- rep(xrange, 2)
  mytype <- factor(rep(c("E6", "E12"), each=2), levels=c("AUROC", "AUPR", "F1", "E6", "E12"))
  dummy <- tibble(Id="progesa-female-lmm", value=xrange, low=xrange, high=xrange, type=mytype)  # Dummy table to use the same x-axis limits for E6 and E12
  #print(dummy)

  #key_size <- grid::convertX(theme_get()$legend.key.size, "mm", valueOnly=TRUE)
  #print(key_size)
  key_size <- 1.8   # This results in box of size 6 mm
  #key_size <- unit(6.096, "mm")
  #key_size <- unit(1.2, "lines")
  #grid::convertX(grid::unit(1.2, "lines"), "mm")
  g <- df %>% ggplot(aes(y=Id, x=value, xmin=low, xmax=high)) + 
    ggstance::geom_pointrangeh(aes(colour=sex, shape=model)) + 
    #geom_pointrange(aes(colour=sex, fill=sex, shape=model), key_glyph = draw_key_rect) + 
    labs(x="Value", y="Data and method", colour="Sex", shape="Model") + 
    #scale_colour_manual(breaks = c("male", "female", "both"), values=c("yellow", "brown", "black"), aesthetics = c("color", "fill", "shape")) +
    scale_colour_discrete(breaks = c("male", "female", "both")) +
    scale_shape_discrete(breaks=c("lmm", "dlmm", "dt", "rf")) +
    #scale_colour_manual(values = c("female" = 1, "male" = 2, "both" = 3)) +
    guides(colour = guide_legend(override.aes = list(shape = 15, linetype=0, size=key_size))) +
    #guides(colour = guide_legend(override.aes = list(shape = 15, linetype=0))) +
    #theme(legend.key.size = theme_get()$legend.key.size) +
    geom_blank(data=dummy) +
    facet_wrap("type", scales="free_x")
  #ggsave(filename="~/FRCBS/results-for-eba-paper/forest_plot.pdf", plot=g, dpi=600, units="mm", width=180)
  return(g)
}

@@

```{r}
g <- create_performance_forest_plot()
if (save_figs)
  ggsave(filename="performance_forest_plot.pdf", title="Performance forest plot", path=fig_path, plot=g, dpi=600, units="mm", width=180)
g
```
## Calibration plots

@@

>>>>

  plot <- result %>% 
    ggplot() +     
    geom_vline(aes(xintercept=0), color="lightgray") +
    geom_rect(data=combined_variables %>% filter(stripe==1), 
              mapping=aes(ymax = as.numeric(Pretty) + 0.5,
                  ymin = as.numeric(Pretty) - 0.5),
                  #xmin = -Inf,
                  #xmax = Inf,
                  fill = "gray", xmin=-Inf, xmax=Inf, alpha = 0.1, show.legend = FALSE, colour=NA) +
    ggstance::geom_pointrangeh(aes(y=Pretty, x=mean, xmin=low, xmax=high, color=gender),
                               position=position_dodge2(width=0.8, padding=0.1), size=0.25) + # ggstance is required to make legend keys horizontal
    labs(title="", #title="Effects sizes of variables on Hb prediction",
         x="Regression coefficient", y=NULL, colour="Sex") +
    scale_colour_discrete(drop=FALSE, breaks = c("male", "female")) + 
    scale_y_discrete() + # !!!!! This is important. Solves the problem with position_dodge2 and the order of rect and pointrange geoms !!!!!!
                         # Otherwise following error results: Error: Discrete value supplied to continuous scale
    #guides(colour = guide_legend(reverse = TRUE)) + # Reverse the order of keys in the legend
    #theme_gray(base_size = base_size) + theme(legend.position = "bottom", legend.direction = "horizontal")
    theme_classic(base_size = base_size) + theme(legend.position = "bottom", legend.direction = "horizontal")
  return(list(plot=plot, cis=cis))
}

@@

  varimp.plot <- var_imp %>% 
    ggplot() + 
    geom_rect(data=combined_variables %>% filter(stripe==1),
              mapping=aes(ymax = as.numeric(Pretty) + 0.5,
                          ymin = as.numeric(Pretty) - 0.5),
              #xmin = -Inf,
              #xmax = Inf,
              #fill = stripe), 
              xmin=-Inf, xmax=Inf, alpha = 0.1, colour=NA, fill="gray", show.legend = FALSE, inherit.aes = FALSE) +
    ggstance::geom_pointrangeh(aes(y=Pretty, x=Importance, xmin=0, xmax=Importance, colour = gender), # ggstance is required to make legend keys horizontal
                               position=position_dodge2(width=0.8, padding=0.1), size=0.25, inherit.aes = FALSE) +   
                               #colour = hue_pal()(3)[2], size=0.25) +
    labs(y=NULL, colour="Sex") + 
    guides(y="none") +
    scale_colour_discrete(drop=FALSE, breaks=c("both")) +
    scale_y_discrete() + # !!!!! This is important. Solves the problem with position_dodge2 and the order of rect and pointrange geoms !!!!!!
                         # Otherwise following error results: Error: Discrete value supplied to continuous scale
    #theme_gray(base_size = base_size)
    theme_classic(base_size = base_size) + theme(legend.position = "bottom", legend.direction = "horizontal")

@@

```{r}
if (save_figs) {
    filename <- paste(fig_path, "effect_size_importance.pdf", sep="/")
} else {
    filename <- NULL
}
create_forest_importance_plot(male_posterior, female_posterior, variables, filename=filename)
```

## Alternative effect sizes and importances of variables

@@

  var_imp <- varimp %>%
    mutate(Pretty=fct_rev(Pretty),
           mean=Importance,
           low=ifelse(is.na(Importance), NA, 0),
           high=Importance)

  final <- bind_rows(`Regression coefficient`=result, Importance=var_imp, .id="method") %>%
    mutate(method=factor(method, levels=c("Regression coefficient", "Importance"))) %>%
    select(-Importance, -Variable, stripe)

  print(combined_variables)

  plot <- final %>% 
    ggplot() +     
    geom_vline(aes(xintercept=0), color="lightgray") +
    # ggstance is required to make legend keys horizontal
    geom_rect(data=combined_variables %>% filter(stripe==1),
              mapping=aes(ymax = as.numeric(Pretty) + 0.5,
                          ymin = as.numeric(Pretty) - 0.5),
              fill = "gray", xmin=-Inf, xmax=Inf, alpha = 0.1, show.legend = FALSE, colour=NA, inherit.aes = FALSE) +
    ggstance::geom_pointrangeh(mapping=aes(y=Pretty, x=mean, xmin=low, xmax=high, color=gender),
                               position=position_dodge2(width=0.8, padding=0.1), 
                               size=0.25,  
                               inherit.aes = FALSE) +
    labs(title=NULL, x=NULL, y=NULL, colour="Sex") +
    scale_colour_discrete(drop=FALSE, breaks = c("male", "female", "both")) + 
    scale_y_discrete() + # !!!!! This is important. Solves the problem with position_dodge2 and the order of rect and pointrange geoms !!!!!!
                         # Otherwise following error results: Error: Discrete value supplied to continuous scale
    facet_wrap("method", scales="free_x") +
    theme_classic(base_size = base_size) + 
    theme(legend.position = "bottom", legend.direction = "horizontal", strip.background = element_rect(fill = NULL, color = "white", size = 1))

@@

```{r}
res <- alternative_create_forest_importance_plot(male_posterior, female_posterior, variables, filename=filename)
g <- res$plot
final <- res$final
if (save_figs)
  ggsave(filename="alternative_effect_size_importance.pdf", title="Performance forest plot", path=fig_path, plot=g, dpi=600, units="mm", width=180)
g
```

@@

get_threshold_df <- function() {
  t6 <- tibble(
      Id = factor(ids),
      type = factor("Threshold"),
      value = c(0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.2, 0.1, 0.4, 0.6))
  t12 <- tibble(
      Id = factor(ids),
      type = factor("Threshold"),
      value = c(0.2, 0.2, 0.1, 0.1, 0.1, 0.2, 0.3, 0.3, 0.8, 0.8)
  )
  t <- bind_rows(t6, t12)
  t <- t %>%
    mutate(low=value, 
           high=value,
           sex=factor(str_split_fixed(Id, "-", 3)[,2], levels=c("female", "both", "male")),
           #model=factor(str_split_fixed(Id, "-", 3)[,3])
           model=rep(c("t6", "t12"), each=length(ids))
           )
  return(t)
}

@@

  thresholds <- get_threshold_df()
  df <- bind_rows(df, thresholds)
  df <- df %>% 
    mutate(model = factor(model, levels=c("lmm", "dlmm", "dt", "rf", "t6", "t12")))
  # Below we use a dummy table and geom_blank to have the same xlimits for E6 and E12 panels
  xrange <- as.numeric(df %>% filter(type %in% c("E6", "E12")) %>% summarise(low=min(low), high=max(high))) # range of x-axis for E6 and E12
  xrange <- rep(xrange, 2)
  mytype <- factor(rep(c("E6", "E12"), each=2), levels=c("AUROC", "AUPR", "F1", "E6", "E12"))
  dummy <- tibble(Id="progesa-female-lmm", value=xrange, low=xrange, high=xrange, type=mytype)  # Dummy table to use the same x-axis limits for E6 and E12
  #print(dummy)

  #key_size <- grid::convertX(theme_get()$legend.key.size, "mm", valueOnly=TRUE)
  #print(key_size)
  key_size <- 1.8   # This results in box of size 6 mm
  #key_size <- unit(6.096, "mm")
  #key_size <- unit(1.2, "lines")
  #grid::convertX(grid::unit(1.2, "lines"), "mm")
  g <- df %>% ggplot(aes(y=Id, x=value, xmin=low, xmax=high)) + 
    ggstance::geom_pointrangeh(aes(colour=sex, shape=model)) + 
    #geom_pointrange(aes(colour=sex, fill=sex, shape=model), key_glyph = draw_key_rect) + 
    labs(x="Value", y="Data and method", colour="Sex", shape="Model") + 
    #scale_colour_manual(breaks = c("male", "female", "both"), values=c("yellow", "brown", "black"), aesthetics = c("color", "fill", "shape")) +
    scale_colour_discrete(breaks = c("male", "female", "both")) +
    scale_shape_manual(breaks=c("lmm", "dlmm", "dt", "rf"), values=c("square", "circle", "triangle", "diamond", "cross", "asterisk")) +
    #scale_colour_manual(values = c("female" = 1, "male" = 2, "both" = 3)) +
    guides(colour = guide_legend(override.aes = list(shape = 15, linetype=0, size=key_size))) +
    #guides(colour = guide_legend(override.aes = list(shape = 15, linetype=0))) +
    #theme(legend.key.size = theme_get()$legend.key.size) +
    geom_blank(data=dummy) +
    facet_wrap("type", scales="free_x")
  #ggsave(filename="~/FRCBS/results-for-eba-paper/forest_plot.pdf", plot=g, dpi=600, units="mm", width=180)
  return(list(g=g, df=df))
}

@@

```{r}
res <- create_performance_forest_plot()
g <- res$g
df <- res$df
if (save_figs)
  ggsave(filename="performance_forest_plot.pdf", title="Performance forest plot", path=fig_path, plot=g, dpi=600, units="mm", width=180)
g
```

## Calibration plots

@@

```{r, eval=FALSE}
jarkko_test <- function(male_posterior, female_posterior, variables,
                                           filename=NULL,
                                           width = 180,  # width of the combined figure in millimetres
                                          base_size = 11
) {
  rrfFit_roc <- load_single("~/FRCBS/interval_prediction/results/rrfFit_roc.rdata")

  var_imp <- varimp %>%
    mutate(Pretty=fct_rev(Pretty),
           mean=Importance,
           low=ifelse(is.na(Importance), NA, 0),
           high=Importance)

  final <- bind_rows(`Regression coefficient`=result, Importance=var_imp, .id="method") %>%
    mutate(method=factor(method, levels=c("Regression coefficient", "Importance"))) %>%
    select(-Importance, -Variable, stripe)

#   final <- tibble(
#             method = c("Regression coefficient",
#                        "Regression coefficient","Importance"),
#             Pretty = factor(c("Days to previous full blood donation","Days to previous full blood donation",
#                        "Days to previous full blood donation")),
#             #gender = c("male", "female", "both"),
#               mean = c(0.0152091681221149,
#                        0.0445741103338152,0.0255604042298423),
#                low = c(0.00691102953212081, 0.0353192308283623, 0),
#               high = c(0.0240855557569598,
#                        0.0537929726515129,0.0255604042298423)
# )

  #combined_variables <- tibble(Pretty=factor("Days to previous full blood donation"), stripe=factor(1))

  print(final)         
  print(combined_variables)
  vars <- c("Days to previous full blood donation", "Age", "Previous donation deferred", "Year")
  plot <- final %>% #filter(Pretty %in% vars) %>%
    ggplot(aes(color=gender)) +
    geom_rect(data=combined_variables %>% filter(stripe==1),
              mapping=aes(#ymax = as.numeric(Pretty) + 0.5,
                          #ymin = as.numeric(Pretty) - 0.5
                          ymax = as.numeric(Pretty) + 0.5,
                          ymin = as.numeric(Pretty) - 0.5),
              fill = "blue", 
              #position="identity",
              xmin=-Inf, xmax=Inf, alpha = 0.1, inherit.aes = FALSE) +
    ggstance::geom_pointrangeh(#data=final, 
                               mapping=aes(y=Pretty, x=mean, xmin=low, xmax=high),
                               #position=position_dodge2(0.8)
                               position=position_dodge2(width=0.8, padding=0.1)
                               ) +
    scale_y_discrete() +
    facet_wrap("method", scales = "free_x") +
    #coord_flip() +
    theme_classic(base_size = base_size)

  return(list(plot=plot, final=final))
}

res <- jarkko_test(male_posterior, female_posterior, variables, filename=filename)
final <- res$final
g <- res$plot
g
```
